---
title: "Project 1 - Financial Data API"
author: "Spencer Williams"
date: "6/25/2023"
output: 
  github_document:
    toc: true
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages Used for Financial Data

Here is a list of packages I will be using to help me create functions and manipulate the data throughout my project.

1. **jsonlite** - This is an API interaction package that will allow me to connect to the Financial database.   It will also allow me to get this data into a nice data frame.

2. **tidyverse** - This package contains `dplyr` and `tidyr` that will let me manipulate and reshape the data frames. I can add/combine variables and dataframes while also creating plots.

3. **rmarkdown** - This package will be used to render my data from Rstudio to Github in the correct repos.

## Loading in the correct packages

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Loading in Data Packages
library(httr)
library(dplyr)
library(jsonlite)
library(rmarkdown)
library(tidyverse)
```

## Convert to Numeric

Creating a function to convert character vectors to numeric if it is a vector of numerical data.

```{r convertToNumeric, include=FALSE}
convertToNumeric <- function(vector){
  # Converting Data to Numeric
  if (any(is.na(suppressWarnings(as.numeric(vector))) == TRUE)){
    output <- vector
  }
  else {
    output <- as.numeric(vector)
  }
  # Return
  return(output)
}
```

# Functions

## Tickers

Creating a function to output data based on the name of the stock. The data contains ticker symbols such as stocks/equities, indices, forex, and crypto.

```{r tickers}
tickers <- function(stock="all"){
  # Get data from API
  outputAPI <- fromJSON("https://api.polygon.io/v3/reference/tickers?active=true&apiKey=BnFt9sqaLWehzloLkpMtEeOG4WjAqkUc")
  # Results tab in API
  output <- outputAPI$results
  # See if the stock matches one of the names in the ticker colunn
  if(stock !="all") {
    if(stock %in% output$ticker) {
      output <- output %>% filter(ticker == stock)
    }
    # See if the stock matches one of the names in the name column
    else if (stock %in% output$name) {
      output <- output %>% filter(name == stock)
    }
    # Error Message
    else {
      message <- paste("Error: Stock was not found in the ticker or name column of dataset.")
      stop(message)
    }
  }
    else {
  
    }
  # Create Data Frame
  output <- suppressMessages(as.data.frame(lapply(output, convertToNumeric)))
  # Return
  return(output)
}
```

## Grouped Daily January

The Grouped Daily function allows the user to get a daily open, high, low, and closing price for the entire stock/equities markets. In this function, I set up data from January 9, 2023.

```{r groupedDailyJan}
groupedDailyJan <- function(stock="all"){
  # Get data from API
  outputAPI <- fromJSON("https://api.polygon.io/v2/aggs/grouped/locale/us/market/stocks/2023-01-09?adjusted=true&apiKey=BnFt9sqaLWehzloLkpMtEeOG4WjAqkUc")
  # Results tab in API
  output <- outputAPI$results
  # See if stock matches one of the names in the column T
  if(stock !="all") {
    if(stock %in% output$T) {
      output <- output %>% filter(T == stock)
    }
    # Error Message
    else {
      message <- paste("Error: Stock was not found in the ticker or name column of dataset.")
      stop(message)
    }
  }
    else {
  
    }
  # Create Data Frame
  output <- suppressMessages(as.data.frame(lapply(output, convertToNumeric)))
  # Return
  return(output)
}
```

## Grouped Daily March

The Grouped Daily function allows the user to get a daily open, high, low, and closing price for the entire stock/equities markets. In this function, I set up data from March 9, 2023.

```{r groupedDailyMar}
groupedDailyMar <- function(stock="all"){
  # Get Data from API
  outputAPI <- fromJSON("https://api.polygon.io/v2/aggs/grouped/locale/us/market/stocks/2023-03-09?adjusted=true&apiKey=BnFt9sqaLWehzloLkpMtEeOG4WjAqkUc")
  # Results tab in API
  output <- outputAPI$results
  # See if the stock matches one of the names in the column T
  if(stock !="all") {
    if(stock %in% output$T) {
      output <- output %>% filter(T == stock)
    }
    # Error Message
    else {
      message <- paste("Error: Stock was not found in the ticker or name column of dataset.")
      stop(message)
    }
  }
    else {
  
    }
  # Create Data Frame
  output <- suppressMessages(as.data.frame(lapply(output, convertToNumeric)))
  # Return
  return(output)
}
```

## Stock Splits V3

Stock Splits V3 will give the user a list of historical stock splits, including the ticker symbol, the execution date, and the factors of the split ratio.

```{r stockSplits}
stockSplits <- function(stock="all"){
  # Get Data from API
  outputAPI <- fromJSON("https://api.polygon.io/v3/reference/splits?apiKey=BnFt9sqaLWehzloLkpMtEeOG4WjAqkUc")
  # Results tab in the API
  output <- outputAPI$results
  # See if the stock matches one of the names in the ticker column
  if(stock !="all") {
    if(stock %in% output$ticker) {
      output <- output %>% filter(ticker == stock)
    }
    # Error Messages
    else {
      message <- paste("Error: Stock has not been split recently or invalid stock.")
      stop(message)
    }
  }
    else {
  
    }
  # Create Data Frame
  output <- suppressMessages(as.data.frame(lapply(output, convertToNumeric)))
  # Return
  return(output)
}
```

## Ticker Types

Ticker Types is the different type of stocks that are being kept in the data system that I am using. It has information such as the code or ticker, the full description name, asset class, and the country.

```{r tickerTypes}
tickerTypes <- function(stock="all"){
  # Get Data from API
  outputAPI <- fromJSON("https://api.polygon.io/v3/reference/tickers/types?apiKey=BnFt9sqaLWehzloLkpMtEeOG4WjAqkUc")
  # Results tab in API
  output <- outputAPI$results
  # See if the stock matches one of the names in the code column
  if(stock !="all") {
    if(stock %in% output$code) {
      output <- output %>% filter(code == stock)
    }
    # See if the stock matches one of the names in the descriptions column
    else if (stock %in% output$description) {
      output <- output %>% filter(description == stock)
    }
    # Error Message
    else {
      message <- paste("Error: Stock was not found in the ticker or name column of dataset.")
      stop(message)
    }
  }
    else {
  
    }
  # Create Data Frame
  output <- suppressMessages(as.data.frame(lapply(output, convertToNumeric)))
  # Return
  return(output)
}
```

## Dividend V3

This function will give a list of historical cash dividends, including the ticker symbol, declaration date, ex-dividend date, record date, pay date, frequency, and amount for CDs in the database.

```{r dividendsV3}
dividendV3 <- function(stock="all"){
  # Get Data from API
  outputAPI <- fromJSON("https://api.polygon.io/v3/reference/dividends?apiKey=BnFt9sqaLWehzloLkpMtEeOG4WjAqkUc")
  # Results tab in API
  output <- outputAPI$results
  # See if Stock matches one of the names in the ticker column
  if(stock !="all") {
    if(stock %in% output$ticker) {
      output <- output %>% filter(ticker == stock)
    }
    # Error Message
    else {
      message <- paste("Error: Stock was not found in the ticker or name column of dataset.")
      stop(message)
    }
  }
    else {
  
    }
  # Create Data Frame
  output <- suppressMessages(as.data.frame(lapply(output, convertToNumeric)))
  # Return
  return(output)
}
```

## Financial API Wrapper Function

```{r financialAPI}
FinancialAPI <- function(func, ...){
  # List of Functions Created
  if (func == "tickers"){
    output <- tickers(...)
  }
  else if (func == "groupedDailyJan"){
    output <- groupedDailyJan(...)
  }
  else if (func == "groupedDailyMar"){
    output <- groupedDailyMar(...)
  }
  else if (func == "stockSplits"){
    output <- stockSplits(...)
  }
  else if (func == "tickerTypes"){
    output <- tickerTypes(...)
  }
  else if (func == "dividendV3"){
    output <- dividendV3(...)
  }
  # Error Message
  else {
    stop("ERROR: Argument for function is not valid!")
  }
  # Return
  return(output)
}
```

# Data Exploration

I am going to create a new *Month* variable for the `groupedDailyJan` and `groupedDailyMar` datasets so that i can combine the two. I am then going to create one more variable, *Difference*, that will calculate the difference between the *Open* and *Close* price.

```{r combination, Include = FALSE}
Jan <- FinancialAPI("groupedDailyJan")
Mar <- FinancialAPI("groupedDailyMar")

# Renaming Column Names
colnames(Jan) <- c("Symbol", "Volume", "Volume_Weight", "Open", "Close", "High", "Low", "Timestamp", "Transactions")
colnames(Mar) <- c("Symbol", "Volume", "Volume_Weight", "Open", "Close", "High", "Low", "Timestamp", "Transactions")

# Creating Month and Difference Variable
Jan <- mutate(Jan, Month = "January", Difference = (Open - Close))
Mar <- mutate(Mar, Month = "March", Difference = (Open - Close))

# Joining Datasets
fullGroupedData <- full_join(Jan, Mar)
head(fullGroupedData)
```

## Contingency Tables

I am creating a `Returns` variable that will tell us whether the stock increased in Price throughout the day. I am then going to create a table to see if January 9th or March 9th would have been a better day for investors. Then, I am going to create a `Twenty` variable to see if stocks were higher or lower than this value at their opening price.

```{r table}
Returns <- vector()
for (i in seq_len(nrow(fullGroupedData))) {
  if(fullGroupedData$Difference[i] >= 0) {
    Returns[i] <- "Positive"
  } else if (fullGroupedData$Difference[i] <= 0) {
    Returns[i] <- "Negative"
  } else {
    Returns[i] <- "Error"
  }
}
fullGroupedData$Returns <- Returns

knitr::kable(table(fullGroupedData$Month, fullGroupedData$Returns))
```

Based on the table, investors had a relatively good day on January 9th and March 9th. March was a very good day for investors with 9,284 stocks being positive out of the total 10,844. January had a respectable 6,668 increased stocks out of the possible 10,953.

```{r contingency2, include=TRUE}
Twenty <- vector()
for (i in seq_len(nrow(fullGroupedData))) {
  if(fullGroupedData$Open[i] >= 20) {
    Twenty[i] <- "Higher"
  } else if (fullGroupedData$Open[i] <= 20) {
    Twenty[i] <- "Lower"
  } else {
    Twenty[i] <- "Error"
  }
}
fullGroupedData$Twenty <- Twenty

knitr::kable(table(fullGroupedData$Month, fullGroupedData$Twenty))
```

Based on this contingency table comparing the opening price of each stock to the month variable, a 20 dollar stock seems to be around the median. In January, 5,316 stocks were above an opening price of 20 dollars while 5,637 stocks were below this price. In March, these totals were 5,418 above and 5,426 below the opening price of 20 dollars